---
title: "The Exponential Distribution"
output: pdf_document
---
## Distribution of Sample Means  
### Lambda = 0.2, 40 samples, 2000 trials  

```{r echo=FALSE}  
library(ggplot2) 
```
```{r}
set.seed(314159)  
lambda <- 0.2;    number.samples <- 40;     nosim <- 2000
means   <- apply(matrix(rexp(number.samples*nosim, lambda) , nosim), 1, mean)
mean.of.sample.means     <- mean(means)
theoretical.mean         <- 1/lambda
variance.of.sample.means <- var(means)
theoretical.variance     <- (1/(number.samples*lambda**2))
```
The theoretical mean is `r theoretical.mean` while the mean of the sample means is `r mean.of.sample.means`   
The theoretical variance is `r theoretical.variance` while the variance of sample means is `r variance.of.sample.means`

```{r histogramOfMeans, fig.width=5.5, fig.height=3.5, fig.cap="Histogram of Sample Means", warning=FALSE} 
 
g <- ggplot(data = data.frame(means), aes(x = means)) +
  geom_histogram(aes(y = ..density..), binwidth= 0.1, fill= 'lightblue', colour='black') +
  geom_density(colour = "blue")+   stat_function(fun = dnorm, colour = "red", size = 1,
             args = list(mean = 5,sd = 1/(lambda)*(1/sqrt(number.samples))))
```

```{r echo=FALSE ,  fig.width=5.5, fig.height=3.5, fig.cap="Histogram of Sample Means", warning=FALSE}
 g + xlab("Means") + 
    ylab("") + 
    ggtitle("Means of Samples from Exponential Distribution") +  
    annotate("text", x = c(7.4,7.5),  y = c(0.5,0.55), size = 3,
               label =  c("Uniform Density", "Density of samples")) +
    annotate("segment", x = c(6.2,6.2), xend = c(6.4,6.4), y  = c(0.5,0.55), 
          yend = c(0.5,0.55),colour = c("red", "blue"))
```

```{r}
shapiro <- as.character( shapiro.test(means)$p.value)
```
Visually, the distribution of sample means comes close to a normal distribution(red), which is confirmed by the results (p.value = ```r shapiro``` ) of the Shapiro-Wilk test.



## Testing Confidence Levels

### Lambda 0.1 to 100.0, 40 samples, 2000 trials

```{r}
 
Lambda.Value <- seq(0.1, 100.0, by = .25);

Coverage <- sapply(Lambda.Value, function(lambda){
  runs   <- matrix(rexp(number.samples*nosim,lambda) , nosim)    
  lhats  <- apply(runs , 1, mean)
  lsd    <- apply(runs,1,sd)
  ll     <- lhats - qnorm(.975) * lsd /sqrt(number.samples)
  ul     <- lhats + qnorm(.975) * lsd /sqrt(number.samples)
  mean(ll < 1/lambda & ul > 1/lambda) })

```
 
 
```{r plotOfCoverage}

ggplot(data.frame(Lambda.Value, Coverage), aes(x = Lambda.Value, y = Coverage)) +  
    geom_line(size = .25) + geom_hline(yintercept = 0.95) + 
    geom_hline(yintercept = 0.925, colour='navy', linetype=2)+
    ylim(.85, 1.0) +  xlab("Lambda") +  
    ggtitle("Coverage of 95% intervals")

```

**Coverage seems to be about 92.5% for the 95% intervals for sample sizes of 40.**  
**The value of Lambda does not seem to affect the coverage.**

[Link to Github](https://github.com/raydante/Statistical_Inference) with the .Rmd file that produced this document.

